---
title: "Gapminder Analysis"
author: "Mar Schmidt"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: yes
      toc_float: yes
      code_folding: hide
      theme: spacelab
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the first line of my document. In this document, I will analyze how life expectancy has changed over time.

# Reading in Gapminder Data

```{r reading-in-data}
gapminder_df <- read.csv("data/gapminder_data.csv")

getwd() 
```

## A smaller header

### An even smaller header

##### A very small header

-   hyphen list
-   hyphen list

Numbered lists:

1.  Number one\
2.  Number two\
3.  Number three

Test if we need a space\
\* test list 1 \* test list 2

We make text **bold with two asterisks** or *italic with one asterisk.*

We can also insert greek letters $\mu$ or $\Delta$

We can make a [hyperlink](https://pad.carpentries.org/Spring2024-BIOMI6300)

**This is bold.**

# Calculating summary statistics

```{r calculating-median-pop, echo = FALSE}
med_pop <- median(gapminder_df$pop)
num_obs <- nrow(gapminder_df)
```

As shown above, the median population was 7023596.

Over `r num_obs` observations, the median population was `r round(med_pop/10e6, digits = 3)`.

# Plotting Life Expectancy over the Years

```{r line-graph, warning = FALSE}
library(ggplot2)

ggplot(gapminder_df, 
       aes(x = year, y = lifeExp, group = country, color = continent)) + 
  geom_line() + 
  scale_y_log10() + 
  scale_x_continuous(limits = c(1997, 2007)) + 
  theme_bw()
```

# Plotting Population distribution

```{r histogram, message = FALSE, fig.width=3, fig.height=3}
ggplot(gapminder_df, aes(x = pop)) + 
  geom_histogram()
```

A table:

| Column 1        | Column 2 | Column 3 |
|-----------------|------------|------|
| Stuff here      | more stuff | wooo |
| blah            | blah       | blah |
| bla             | blah       | blah |


we're adding a new line because we're now working on git. 

